---
title: "p8105_hw1_lg3239"
author: "Landi Guo"
date: "2022-09-20"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE}
library(ggplot2)
library(tidyverse)
```

## Problem 1
```{r}
# load in the penguins dataset
data("penguins", package = "palmerpenguins")
```

Details of the dataset:

```{r}
skimr::skim(penguins)
```

This dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns, which consisting of 3 factor columns and 5 numeric columns. The column names are `r names(penguins)`. Only species, island, and year do not contain missing values. The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.

```{r}
# ggplot() is used to load the penguins dataset with bill length on x-axis and flipper length on y-axis. The color is set to correspond species. geom_point() is used to generate scatter plot, and labs() is used to specify plot title.

bill_flipper_plot = ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "flipper_length_mm vs bill_length_mm")
```

```{r}
# save the scatter plot above locally as a pdf
ggsave(filename = "hw1_plot.pdf", plot = bill_flipper_plot)
```

## Problem 2
Using `tibble()` to create a dataframe according to the requirements.
```{r}
p2_df = 
  tibble(
    norm = rnorm(n = 10),
    logical = norm > 0,
    char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
    factor = factor(
      c("low", "high", "medium", "medium", "low", "high", "low", "high", "medium", "high"),
      levels = c("high", "medium", "low")
    )
  )

p2_df
```

```{r}
mean(p2_df$norm)
mean(p2_df$logical)
mean(p2_df$char)
mean(p2_df$factor)
```
Only taking the means of `norm` and `logical` work, the means of `char` and `factor` does not work.

```{r, results='hide'}
as.numeric(p2_df$logical)
as.numeric(p2_df$char)
as.numeric(p2_df$factor)
```
The column of `logical` gets converted to numeric values of 0 and 1, indicating `FALSE` and `TRUE`. The column of `char` becomes `NA`s. The column of `factor` becomes values of 1, 2, 3 corresponding to `high`, `medium`, `low`.
